### 2019-04-22 initial/notes.tex
update notes

@@

We zoom in on Cassini State $4$, which has $\cos \theta \approx 0, \phi = 0$. In
particular, by using equations of motion
\begin{align}
    \pd{\phi}{t} &= \cos \theta - \eta\p*{\cos I + \sin I \cot \theta \cos \phi}
        ,\\
    \pd{\theta}{t} &= -\eta \sin I \sin \phi - \s*{\epsilon \sin \theta},
\end{align}
we can compute $\theta_4 \approx -\frac{\pi}{2} +\frac{\eta \sin I}{1 + \eta
\cos I}, \phi_4 = 0$ in the $\epsilon = 0$ limit (the other solution is
$\theta_2 \approx +\frac{\pi}{2} - (\dots)$). Indeed the signs are correct:
near $\theta \approx \frac{\pi}{2}, \phi \approx 0$ we have EOM
$\dot{\delta\phi}_4 \approx -\delta \theta_4$ while $\dot{\delta \theta_4}
\approx -\eta \sin I \delta \phi_4$ which is unstable.

@@

>>>>

\subsection{Separatrix Area}

We can estimate the area enclosed by the separatrix, as shown in
\autoref{fig:1contours}. Note that the separatrix joins Cassini State 4 to its
$+ 2\pi$ image.
\begin{figure}[t]
    \centering
    \includegraphics[width=0.6\textwidth]{plots/1contours.png}
    \caption{Separatrix for various values of $\eta$.}\label{fig:1contours}
\end{figure}

We notate $\mu = \cos\theta$; note that CS4 is $\mu_4 \approx \frac{\eta \cos
I}{1 + \eta \sin I} \approx \eta \cos I$. Setting the Hamiltonian equal to its
value at CS4 gives
\begin{align*}
    H_4 &\equiv H\p*{\mu_4, \phi_4}
        \approx \frac{\mu_4^2}{2} - \eta \mu_4 \cos I + \eta \sin I,\\
        &= -\eta^2\cos^2 I + \eta \sin I,\\
    H(\mu_{sep}, \phi_{sep})
        &= H_4 = \eta \sin I \cos \phi + \frac{\mu^2}{2} - \eta \mu \cos I
            + \mathcal{O}(\eta^3),\\
    0 &\approx \frac{\mu_{sep}^2}{2} - \eta \mu_{sep} \cos I
        - \eta \sin I\p*{1 - \cos \phi_{sep}} + \eta^2\cos^2I,\\
    \mu_{sep}\p*{\phi} &\approx \sqrt{2\eta \sin I\p*{1 - \cos \phi}}
        + \mathcal{O}(\eta).
\end{align*}

We can then easily compute the area enclosed by the separatrix
\begin{align}
    A_{sep} &= \int\limits_0^{2\pi}2\mu_{sep}\;\mathrm{d}\phi,\nonumber\\
        &\approx 16\sqrt{\eta \sin I}.
\end{align}
For $\eta = 0.1, I = 20^\circ$, this predicts $\frac{A_{sep}}{A_{T}} \approx
0.235$, which is pretty close to my numerically calculated $\frac{A_{sep}}{A_T}
= 0.229$.

@@

We zoom in on Cassini State $4$, which has $\cos \theta \approx 0, \phi = 0$. In
particular, by using equations of motion
\begin{align}
    \pd{\phi}{t} &= \cos \theta - \eta\p*{\cos I + \sin I \cot \theta \cos \phi}
        ,\\
    \pd{\theta}{t} &= -\eta \sin I \sin \phi - \s*{\epsilon \sin \theta},
\end{align}
we can compute $\theta_4 \approx -\frac{\pi}{2} +\frac{\eta \cos I}{1 + \eta
\sin I}, \phi_4 = 0$ in the $\epsilon = 0$ limit (the other solution is
$\theta_2 \approx +\frac{\pi}{2} - (\dots)$). Indeed the signs are correct:
near $\theta \approx \frac{\pi}{2}, \phi \approx 0$ we have EOM
$\dot{\delta\phi}_4 \approx -\delta \theta_4$ while $\dot{\delta \theta_4}
\approx -\eta \sin I \delta \phi_4$ which is unstable.

@@

\subsection{Melnikov Distance}

We notice that the separatrix is a heteroclinic orbit, or a saddle connection,
in the dissipation free problem. Introducing dissipation breaks the saddle
connection by a distance that can be estimated with the Melnikov distance. This
is G\&H Equation 4.5.11 or something:
\begin{align}
    d(t_0) &= \frac{\epsilon M(t_0)}{\abs*{f(q^0(0))}} + \mathcal{O}(\epsilon^2)
        ,\\
    M(t_0) &= \int\limits_{-\infty}^\infty
        \s*{f \times g}_{hetero}\;\mathrm{d}t.
\end{align}
We evaluate the Melnikov integral $M(t_0)$ on the heteroclinic orbit. Note that
since in our problem our perturbation $g$ is time-independent, so too is the
Melnikov integral $M(t_0) = M$.

Let's apply this to the Cassini state Hamiltonian w/ dissipation. We first write
down our EOM in Melnikov form (we use canonical variables $\mu, \phi$):
\begin{align}
    \rd{\hat{s}}{t}
        &= \underbrace{
            \pd{\mathcal{H}}{\mu}\hat{\phi} - \pd{\mathcal{H}}{\phi}\hat{\mu}}_f
            + \epsilon \underbrace{\p*{1 - \mu^2}\hat{\mu}}_g.
\end{align}
Then $f \times g = f_\phi g_\mu = \pd{\mathcal{H}}{\mu}\p*{1 - \mu^2}$. We then
want to integrate this along the heteroclinic orbit. We can make change of
variables
\begin{equation}
    M = \int\limits_0^{2\pi}\pd{\mathcal{H}}{\mu}
        \p*{1 - \mu^2}\p*{\pd{\phi}{t}}^{-1}\;\mathrm{d}\phi.
\end{equation}
But thankfully, $\pd{\mathcal{H}}{\mu} = \pd{\phi}{t}$ in the absence of
dissipation, and so $M = 2\pi\p*{1 - \mu^2} \approx 2\pi\p*{1 - 2\eta \sin I}$.
Thus, the Melnikov distance at point $q^0$, a point on the heteroclinic
orbit of the unperturbed Hamiltonian, is just
\begin{equation}
    d(q^0) = \frac{2\pi \epsilon\p*{1 - 2\eta \sin I}}{\abs*{f(q^0)}}.
\end{equation}
